Bringing the Power of eBPF to Open vSwitch

### 2018-10-21 paper.tex
afxdp performance

\subsection{Performance Evaluation}
When sending and receivng packets using AF\_XDP with OVS, two CPUs show
100\% utilitization:
\begin{itemize}
\item ksoftirqd:
\item ovs-vswitchd:
\end{itemize}
PMD
HugePage
UMEM element list
Batching
2-CPU operation
non-blocking send
non-block receive

>>>>

\subsection{Performance Evaluation}
% testbed

When forwarding packets using AF\_XDP with OVS, two CPUs show up
100\% utilitization:
\begin{itemize}
\item ovs-vswitchd: This is the process keeps doing the send and receive steps
in Figure~\ref{afxdprx,afxdptx}.
\item ksoftirqd: This is the kernel software interrupt thread handling the
incoming packets, triggering XDP program to pass to the XSK, and also processing
transmission.
\end{itemize}

For all our experiments, we use a microbenchmark program, called xdpsock,
as the baseline to compare with our implementation.
xdpsock is an AF\_XDP sample program doing no packet processing, and can be configured
to do simply dropping the packet, rxdrop, or forwarding the packet to the same port as
it receives, l2fwd.  In our testbed, we measured $19Mpps$ for rxdrop and $17Mpps$ for
l2fwd.

Compared to xdpsock, ovs-vswitchd adds on top of it a UMEM memory pool,
and packets parsing, lookup and action execution.
Our initial prototype shows pretty terrible performance, only $0.5Mpps$ rxdrop.
We investigated and our first fix is to enable OVS's PMD-mode,
Poll-Mode-Driver, to the afxdp netdev.
In OVS, a non-pmd mode netdev does packet receiption by putting all the receiving
netdev's file descript together and ovs-vswitch polls them when there is an fd ready
to perform I/O.  As a result, the fd of the XSK is {\em shared} with other fds, and
we observed that the poll system call has pretty high overhead.

Applying OVS's PMD netdev avoids these problems and improving the rxdrop performance
from $0.5Mpss$ to $3Mpps$.  Ideally, receing packets from XSK does not need to context
switch to into kernel as mentioned in steps~\ref{afxdprx}. So an ideal implementation
should show minimal context switches with most of the CPU stays in userspace. 
When enabling PMD netdev for afxdp, ovs-vswitch can be
configured to create a {\em dedicated} thread for the packet I/O.  
In our case, we create a thread for each XSK receive queue which keeps polling
the RX ring for new packets.  Each round of receive polling processes a batch of
packet, up to maximum of 32 packets. In addition, we enable high memory page 1GB to
reduce the page fault overhead.  With the above fixes/optimizations, the rxdrop
performance increases to around $10Mpps$

For every stage of the optimization, we use Linux perf extensively, e.g., perf stat
and perf record/report.  With the above new design, we observe the new bottleneck
is at the UMEM memory pool we introduce to maintain the UMEM freelist, and the
packet metadata allocation, struct dp\_packet, for every receiving packet.
We introduce two major API, umem\_elem\_get(), which gets N free element from the
UMEM freelist, and umem\_elem\_put(), which places back the free UMEM buffer to
the freelist.
producer 
consumer
We implemented three data structures of UMEM memory pool as below
to compare their performance
\begin{itemize}
\item list\_head: Similar to Linux kernel's list\_head, each element contains it
UMEM addr, and a next pointer points to next free element. Elements are sparse in
memory. Consumer always takes the first element, head, and producer also place back
from the head of the list.
\item ptr\_ring: Similar to Linux kernel's ptr\_ring, an array of elements are
allocated in continuous memory region, and pointers (head and tail) point for
Consumer takes elements from the tail and when placing back, producer, puts
them into where head pointer points to.
\item ptr\_stack: Similar to ptr\_ring, but consumer and producer always use
the top few element, thus similar to push pop. Compared

\end{itemize}
% lock
% push/pop
% preallocate array
Since the UMEM

We then observe that the UMEM memory pool implementation ha

PMD
HugePage
UMEM element list
Batching
2-CPU operation
non-blocking send
non-block receive
pre-allocation dp\_packet
rxhash
